Static Analysis of Tail Behaviour with a \\ Generalized Gamma Algebra

### 2022-11-07 design/continuous_nips/continuous.tex
Updates from Overleaf

 \newif\ifisarxiv
 \isarxivfalse

%  \ifisarxiv
 \documentclass[thesis.tex]{subfiles}
%  \usepackage[round]{natbib}
%  \usepackage{fullpage}
%  \usepackage{amsmath}
%  \usepackage{amssymb}
%  \usepackage{amsthm}
%  \usepackage{mathabx}
%  \usepackage{color}
%  \usepackage{cancel}
%  \usepackage{wrapfig}
%  \usepackage{subfigure}
%  \usepackage{caption}
%  \usepackage{xfrac}
%  \usepackage{algorithm}
%  \usepackage{algorithmic}
%  \usepackage{mdframed}
%  \usepackage{hyperref}
%  \usepackage{xcolor}
%  \hypersetup{
%  	colorlinks,
%  	linkcolor={red!40!gray},
%  	citecolor={blue!40!gray},
%  	urlcolor={blue!70!gray}
%  }
%  \usepackage{cleveref}

 \title{Exact expressions for double descent
 and implicit regularization
 \\
 via surrogate random design}
   \author{%
           \textbf{Micha{\l } Derezi\'{n}ski} \\
   Department of Statistics\\
   University of California, Berkeley\\
   \texttt{mderezin@berkeley.edu}\\
   \and
   \textbf{Feynman Liang} \\
   Department of Statistics\\
   University of California, Berkeley\\
   \texttt{feynman@berkeley.edu}
   \and
    \textbf{Michael W. Mahoney}\\
   ICSI and Department of Statistics\\
   University of California, Berkeley\\
   \texttt{mmahoney@stat.berkeley.edu}
 }

 %  \newcommand{\jmlrBlackBox}{\rule{1.5ex}{1.5ex}}
   \providecommand{\BlackBox}{\rule{1.5ex}{1.5ex}}
 %  \newcommand{\jmlrQED}{\hfill\jmlrBlackBox\par\bigskip}
   \renewenvironment{proof}%
   {%
    \par\noindent{\bfseries\upshape Proof\ }%
   }%
 {\hfill\BlackBox\par\bigskip}

@@

\begin{figure}[t]
  \centering
  \begin{subfigure}{0.49\textwidth}%
    \subcaption{Surrogate MSE expressions (Theorem \ref{t:mse}) closely match
    numerical estimates even for non-isotropic
    features. Eigenvalue decay leads to a steeper
    descent curve in the under-determined regime ($n<d$).}
    \includegraphics[width=0.48\textwidth]{../figs/descent-intro}
  \end{subfigure}
  \hfill
  \begin{subfigure}{0.49\textwidth}
    \subcaption{The mean of the estimator $\X^\dagger\y$ exhibits
    shrinkage which closely matches the shrinkage of a
    ridge-regularized least squares optimum (theory lines), as characterized by
    Theorem \ref{t:unbiased}.}
    \includegraphics[width=0.48\textwidth]{../figs/descent-shrinkage}
  \end{subfigure}
  \caption{Illustration of the main results for $d=100$ and
    $\mu=\Nc(\zero,\Sigmab)$ where $\Sigmab$ is diagonal with
    eigenvalues decaying exponentially and scaled so that
    $\tr(\Sigmab^{-1})=d$. We use our surrogate
    formulas to plot (a) the MSE (Theorem \ref{t:mse}) and (b) the norm of the expectation (Theorem
    \ref{t:unbiased}) of the Moore-Penrose estimator (\emph{theory}
    lines), accompanied by the empirical estimates based on the standard
    i.i.d.~design (error bars are three times the standard error of the
    mean). We consider three different condition numbers $\kappa$ of
    $\Sigmab$, with \emph{isotropic} corresponding to $\kappa=1$,
    i.e., $\Sigmab=\I$. We use $\sigma^2=1$ and
    $\w^*=\frac1{\sqrt{d}}\one$.}
  %\vspace{-5mm}
  \label{f:intro}
\end{figure}

@@

\begin{wrapfigure}{r}{\ifisarxiv 0.5\else 0.44\fi\textwidth}
  \ifisarxiv\else \vspace{-.2cm}\fi
  \centering
  \includegraphics[width=.43\textwidth]{../figs/descent-model-log}
  \ifisarxiv\else \vspace{-3mm}\fi
  \caption{
    Surrogate MSE as a function of $d/n$, with $n$
    fixed to $100$ and varying dimension $d$ and condition number
    $\kappa$, for signal-to-noise ratio $\mathrm{SNR}=1$.}
  \label{f:model}
  \ifisarxiv\else  \vspace{-.5cm}\fi
\end{wrapfigure}

@@

% Our work opens up a number of new directions for future research.
% This includes extending our surrogate analysis to the
% misspecified linear regression discussed above.
% Also, it remains open whether the analysis we provided for the mean
% squared error can be reproduced in the context of mean squared
% \emph{prediction} error, which is relevant in many machine learning
% tasks. Finally, while Theorem \ref{t:asymptotic} states that our surrogate
% expressions for the MSE under certain Gaussian designs are
% asymptotically consistent with the multiplicative error rate of
% $O(1/d)$, we believe that this fact extends to
% the setting not covered by the theorem: under-determined regime (i.e.,
% $n<d$) with a non-isotropic Gaussian distribution (i.e., $\Sigmab\neq
% \I$). We break down our analysis into verifying two
% conjectures which are of independent interest to multivariate
% Gaussian analysis.
\ifisarxiv\else
  \section*{Broader Impact}

  While the double descent phenomenon has been empirically observed in a variety
  of applications, mathematical descriptions of it are oftentimes complex and
  inaccesible to non-experts. In contrast, our surrogate design and the
  accompanying closed-form expressions provide an easily computable rule of thumb
  for estimating the generalization error. One important application is
  the high-dimensional (i.e. underdetermined, $n < d$) regime
  experienced by modern machine learning systems where the number of
  parameters vastly exceeds the quantity of
  available data. Our research can be applied here to provide a theoretical
  understanding of the surprising phenomenon where without proper
  regularization \citep{nakkiran2020optimal} more data (i.e. increasing $n$)
  may lead to worse generalization performance.

  Better theoretical understanding of generalization error in the small data
  regime has important societal impact. Through theoretically modeling a system's
  performance, we can build safer systems by better understanding how badly an
  estimator fails and accounting for these failure modes in the system's design.
  Furthermore, the improved understanding of minimum norm solutions performing
  worse with more data offers an appealing trade-off where certain systems can
  both improve their performance and respect the privacy of its users by
  collecting less data.
\fi

\paragraph{Acknowledgements.}
We would like to acknowledge ARO, DARPA, NSF, ONR, and GFSD for providing
partial support of this work. We also thank Zhenyu Liao for pointing
out fruitful connections between our results and the asymptotic
analysis of random matrix resolvents.

>>>>

  \documentclass[thesis.tex]{subfiles}
%  \usepackage[round]{natbib}
%  \usepackage{fullpage}
%  \usepackage{amsmath}
%  \usepackage{amssymb}
%  \usepackage{amsthm}
%  \usepackage{mathabx}
%  \usepackage{color}
%  \usepackage{cancel}
%  \usepackage{wrapfig}
%  \usepackage{subfigure}
%  \usepackage{caption}
%  \usepackage{xfrac}
%  \usepackage{algorithm}
%  \usepackage{algorithmic}
%  \usepackage{mdframed}
%  \usepackage{hyperref}
%  \usepackage{xcolor}
%  \hypersetup{
%  	colorlinks,
%  	linkcolor={red!40!gray},
%  	citecolor={blue!40!gray},
%  	urlcolor={blue!70!gray}
%  }
%  \usepackage{cleveref}

%  \title{Exact expressions for double descent
%  and implicit regularization
%  \\
%  via surrogate random design}
%   \author{%
%           \textbf{Micha{\l } Derezi\'{n}ski} \\
%   Department of Statistics\\
%   University of California, Berkeley\\
%   \texttt{mderezin@berkeley.edu}\\
%   \and
%   \textbf{Feynman Liang} \\
%   Department of Statistics\\
%   University of California, Berkeley\\
%   \texttt{feynman@berkeley.edu}
%   \and
%     \textbf{Michael W. Mahoney}\\
%   ICSI and Department of Statistics\\
%   University of California, Berkeley\\
%   \texttt{mmahoney@stat.berkeley.edu}
%  }

%  %  \newcommand{\jmlrBlackBox}{\rule{1.5ex}{1.5ex}}
%   \providecommand{\BlackBox}{\rule{1.5ex}{1.5ex}}
%  %  \newcommand{\jmlrQED}{\hfill\jmlrBlackBox\par\bigskip}
%   \renewenvironment{proof}%
%   {%
%     \par\noindent{\bfseries\upshape Proof\ }%
%   }%
%  {\hfill\BlackBox\par\bigskip}

@@

\begin{figure}[t]
  \centering
  \begin{subfigure}{0.49\textwidth}%
    \subcaption{Surrogate MSE expressions (Theorem \ref{t:mse}) closely match
    numerical estimates even for non-isotropic
    features. Eigenvalue decay leads to a steeper
    descent curve in the under-determined regime ($n<d$).}
    \includegraphics[width=\textwidth]{../figs/descent-intro}
  \end{subfigure}
  \hfill
  \begin{subfigure}{0.49\textwidth}
    \subcaption{The mean of the estimator $\X^\dagger\y$ exhibits
    shrinkage which closely matches the shrinkage of a
    ridge-regularized least squares optimum (theory lines), as characterized by
    Theorem \ref{t:unbiased}.}
    \includegraphics[width=\textwidth]{../figs/descent-shrinkage}
  \end{subfigure}
  \caption{Illustration of the main results for $d=100$ and
    $\mu=\Nc(\zero,\Sigmab)$ where $\Sigmab$ is diagonal with
    eigenvalues decaying exponentially and scaled so that
    $\tr(\Sigmab^{-1})=d$. We use our surrogate
    formulas to plot (a) the MSE (Theorem \ref{t:mse}) and (b) the norm of the expectation (Theorem
    \ref{t:unbiased}) of the Moore-Penrose estimator (\emph{theory}
    lines), accompanied by the empirical estimates based on the standard
    i.i.d.~design (error bars are three times the standard error of the
    mean). We consider three different condition numbers $\kappa$ of
    $\Sigmab$, with \emph{isotropic} corresponding to $\kappa=1$,
    i.e., $\Sigmab=\I$. We use $\sigma^2=1$ and
    $\w^*=\frac1{\sqrt{d}}\one$.}
  %\vspace{-5mm}
  \label{f:intro}
\end{figure}

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=.5\textwidth]{../figs/descent-model-log}
  \caption{
    Surrogate MSE as a function of $d/n$, with $n$
    fixed to $100$ and varying dimension $d$ and condition number
    $\kappa$, for signal-to-noise ratio $\mathrm{SNR}=1$.}
  \label{f:model}
\end{figure}

@@

% Our work opens up a number of new directions for future research.
% This includes extending our surrogate analysis to the
% misspecified linear regression discussed above.
% Also, it remains open whether the analysis we provided for the mean
% squared error can be reproduced in the context of mean squared
% \emph{prediction} error, which is relevant in many machine learning
% tasks. Finally, while Theorem \ref{t:asymptotic} states that our surrogate
% expressions for the MSE under certain Gaussian designs are
% asymptotically consistent with the multiplicative error rate of
% $O(1/d)$, we believe that this fact extends to
% the setting not covered by the theorem: under-determined regime (i.e.,
% $n<d$) with a non-isotropic Gaussian distribution (i.e., $\Sigmab\neq
% \I$). We break down our analysis into verifying two
% conjectures which are of independent interest to multivariate
% Gaussian analysis.
% \ifisarxiv\else
%   \section*{Broader Impact}

%   While the double descent phenomenon has been empirically observed in a variety
%   of applications, mathematical descriptions of it are oftentimes complex and
%   inaccesible to non-experts. In contrast, our surrogate design and the
%   accompanying closed-form expressions provide an easily computable rule of thumb
%   for estimating the generalization error. One important application is
%   the high-dimensional (i.e. underdetermined, $n < d$) regime
%   experienced by modern machine learning systems where the number of
%   parameters vastly exceeds the quantity of
%   available data. Our research can be applied here to provide a theoretical
%   understanding of the surprising phenomenon where without proper
%   regularization \citep{nakkiran2020optimal} more data (i.e. increasing $n$)
%   may lead to worse generalization performance.

%   Better theoretical understanding of generalization error in the small data
%   regime has important societal impact. Through theoretically modeling a system's
%   performance, we can build safer systems by better understanding how badly an
%   estimator fails and accounting for these failure modes in the system's design.
%   Furthermore, the improved understanding of minimum norm solutions performing
%   worse with more data offers an appealing trade-off where certain systems can
%   both improve their performance and respect the privacy of its users by
%   collecting less data.
% \fi

% \paragraph{Acknowledgements.}
% We would like to acknowledge ARO, DARPA, NSF, ONR, and GFSD for providing
% partial support of this work. We also thank Zhenyu Liao for pointing
% out fruitful connections between our results and the asymptotic
% analysis of random matrix resolvents.

### 2022-11-07 hta/neurips_2022.tex
Updates from Overleaf

\documentclass[thesis.tex]{subfiles}
% \documentclass{article}
\csname remappendix\endcsname

@@

\author{%
  David S.~Hippocampus\thanks{Use footnote for providing further information
    about author (webpage, alternative address)---\emph{not} for acknowledging
    funding agencies.} \\
  Department of Computer Science\\
  Cranberry-Lemon University\\
  Pittsburgh, PA 15213 \\
  \texttt{hippo@cs.cranberry-lemon.edu} \\
  % examples of more authors
  % \And
  % Coauthor \\
  % Affiliation \\
  % Address \\
  % \texttt{email} \\
  % \AND
  % Coauthor \\
  % Affiliation \\
  % Address \\
  % \texttt{email} \\
  % \And
  % Coauthor \\
  % Affiliation \\
  % Address \\
  % \texttt{email} \\
  % \And
  % Coauthor \\
  % Affiliation \\
  % Address \\
  % \texttt{email} \\
}

\sisetup{detect-all = true}

@@

\begin{table}[t]
    \centering
    \bgroup
\def\arraystretch{2.5}
\tolerance=1
\emergencystretch=\maxdimen
\hyphenpenalty=10000
\hbadness=10000
\begin{tabular}{|M{0.15\linewidth} | p{0.75\linewidth}|}
\hline
\centering \multirow{2}{*}{\textbf{Ordering}} & 
$\max\{(\nu_1,\sigma_1,\rho_1),(\nu_2,\sigma_2,\rho_2)\}$ \\ & $\qquad \equiv \begin{cases}
(\nu_1,\sigma_1,\rho_1) & \text{ if } \limsup_{x\to\infty} \frac{x^\nu_1 e^{-\sigma_1 x^{\rho_1}}}{x^\nu_2 e^{-\sigma_2 x^{\rho_2}}} < +\infty \\
(\nu_2,\sigma_2,\rho_2) & \text{ otherwise.}
\end{cases}$ \\ \hline
\centering \multirow{2}{*}{\textbf{Addition}} &  $(\nu_{1},\sigma_{1},\rho_{1})\oplus(\nu_{2},\sigma_{2},\rho_{2})$ \\
 & $\equiv \begin{cases}
\max\{(\nu_{1},\sigma_{1},\rho_{1}),(\nu_{2},\sigma_{2},\rho_{2})\} & \text{ if }\rho_{1}\neq\rho_{2}\text{ or }\rho_{1},\rho_{2}<1\\
\left(\nu_{1}+\nu_{2}+1,\min\{\sigma_{1},\sigma_{2}\},1\right) & \text{ if }\rho_{1}=\rho_{2}=1\\
(\nu_{1}+\nu_{2}+\frac{2-\rho}{2},(\sigma_{1}^{-\frac{1}{\rho-1}}+\sigma_{2}^{-\frac{1}{\rho-1}})^{1-\rho},\rho) & \text{ if }\rho=\rho_{1}=\rho_{2}>1.
\end{cases}$ 
\\ \hline
\centering \textbf{Powers} & $(\nu,\sigma,\rho)^\beta \equiv (\frac{\nu+1}{\beta} - 1,\sigma,\frac{\rho}{\beta})$ for $\beta > 0$ \\\hline
\centering \textbf{Reciprocal*} & $(\nu,\sigma,\rho)^{-1} \equiv \begin{cases} (-\nu-2,\sigma,-\rho) & \text{ if } (\nu + 1)/\rho > 0 \text{ and } \rho \neq 0 \\
\mathcal{R}_2 & \text{ otherwise}
\end{cases}$ \\ \hline
\centering \textbf{Scalar \linebreak Multiplication} & $c(\nu,\sigma,\rho) \equiv (\nu,\sigma/|c|^\rho,\rho)$\\\hline
\multirow{3}{*}{\centering \textbf{Multiplication}} & $(\nu_{1},\sigma_{1},\rho_{1})\otimes(\nu_{2},\sigma_{2},\rho_{2})$ \\
&$\qquad \equiv\begin{cases}
\left(\frac{1}{\mu}\left(\frac{\nu_{1}}{|\rho_{1}|}+\frac{\nu_{2}}{|\rho_{2}|}+\frac{1}{2}\right),\sigma,-\frac{1}{\mu}\right) & \text{ if }\rho_{1},\rho_{2}<0\\
\left(\frac{1}{\mu}\left(\frac{\nu_{1}}{\rho_{1}}+\frac{\nu_{2}}{\rho_{2}}-\frac{1}{2}\right),\sigma,\frac{1}{\mu}\right) & \text{ if }\rho_{1},\rho_{2}>0\\
\mathcal{R}_{|\nu_1|} & \mbox{ if }\rho_{1}\leq0,\rho_{2}>0 \\
\mathcal{R}_{\min\{|\nu_1|,|\nu_2|\}} & \mbox{ if }\rho_{1}=0,\rho_{2}=0
\end{cases}$ \\
&where $\mu=\frac{1}{|\rho_{1}|}+\frac{1}{|\rho_{2}|}=\frac{|\rho_{1}|+|\rho_{2}|}{|\rho_{1}\rho_{2}|}$, $\sigma=\mu(\sigma_{1}|\rho_{1}|)^{\frac{1}{\mu|\rho_{1}|}}(\sigma_{2}|\rho_{2}|)^{\frac{1}{\mu|\rho_{2}|}}$.\\ \hline
\centering \textbf{Product of Densities} & $(\nu_1,\sigma_1,\rho_1)\&(\nu_2,\sigma_2,\rho_2) 
\equiv \begin{cases}
(\nu_{1}+\nu_{2},\sigma_{1},\rho_{1}) & \text{ if }\rho_{1}<\rho_{2}\\
(\nu_{1}+\nu_{2},\sigma_{1}+\sigma_{2},\rho) & \text{ if }\rho=\rho_{1}=\rho_{2}\\
(\nu_{1}+\nu_{2},\sigma_{2},\rho_{2}) & \text{ otherwise.}
\end{cases}$ \\ \hline
\centering \textbf{Functions ($L$-Lipschitz)} & 
$f(X_1,\dots,X_n) \equiv L \max\{X_1,\dots,X_n\}$ \\ \hline
    \end{tabular}
    \vspace{.25cm}
    \caption{\label{tab:gga_operations}Operations on random variables (e.g.$X_1 + X_2$) are 
    viewed as actions on density functions
    (e.g. convolution $(\nu_1, \sigma_1, \rho_1)\oplus (\nu_2, \sigma_2, \rho_2)$) and the tail parameters of the result are analyzed and reported.}
\egroup
\end{table}

@@

\textbf{Variational Inference} The optimization objective is the ELBO $E_q \log \frac{p(X)}{q(X)} \approx \frac{1}{N} \sum_{i=1}^N \log \frac{p(x_i)}{q(x_i)},\quad x_i \sim q$. Here, the density $p$ must also be evaluated so for simplicity
experiments in \cref{tab:vi} use closed-form marginalized densities for targets.
The overall trends also show that GGA yields consistent improvements as measured by both ELBO and importance-weighted
estimates of marginal likelihood and that the difference was greater when the tails of $p(z)$ were heavier.
The $\hat{k}$ diagnostics \citep{yao2018yes} corroborate our findings that variational inference succeeds
($\hat{k} < -1.2$) when a GGA with appropriately matched tails is used and fails ($\hat{k} > 1$) when
Gaussian tails are erroneously imposed.

@@

\begin{figure}[h]
	\centering
	\includegraphics[width=0.48\textwidth]{figures/LogNormal_Bulk.png}
	\includegraphics[width=0.48\textwidth]{figures/LogNormal_Tail.png}
	\caption{Estimation of the log-normal density by tail algebra applied to $V_k$.}
	\label{fig:lognormal}
\end{figure}

@@

Consider the toy example of a Normal target. This case is trivial for
Gaussian based methods and is oftentimes the initialization.
This lack of approximation gap in ADVI is seen in \Cref{fig:normal},
where we also see that GGA achieves similar approximation quality.
This is unsurprising as the GGA approximation in \cref{tab:dist_list}
is also a Normal distribution.

\begin{figure}[H]
	\centering
	\includegraphics[width=.6\textwidth]{figures/de_normal_sum.png}
	\includegraphics[width=.6\textwidth]{figures/vi_normal_sum.png}
	\caption{Density estimation and VI against a known normal target}
	\label{fig:normal}
\end{figure}

@@

Now let $X_{ij} \sim N(0,1)$ and consider $tr X^\top X$.
Such quantities arise in the analysis of random projections.
It is important here to recognize that the power operation $X \mapsto X^2$ is not equivalent to the multiplication operation $X \mapsto X \otimes X$, as multiplication assumes independence.
\begin{figure}[H]
	\centering
	\includegraphics[width=\textwidth]{figures/jl-chi2.png}
	\caption{$5000$ samples of JL matrix trace (blue) vs GGA prediction (yellow)}
	\label{fig:jl_chi2}
\end{figure}

@@

@@

Generalized hyperbolic distribution & $(-\infty,\infty)$ & $\propto e^{\beta(x-\mu)}\frac{K_{\lambda-1/2}(\alpha\sqrt{\delta^{2}+(x-\mu)^{2}})}{(\delta^{2}+(x-\mu)^{2})^{1/4-\lambda/2}}$ & $(\lambda-1,\alpha-\beta,1)$\tabularnewline

@@

Generalized Normal distribution & $(-\infty,\infty)$ & $\frac{\beta}{2\alpha\Gamma(1/\beta)}\exp\left(-\left(\frac{|x-\mu|}{\alpha}\right)^{\beta}\right)$ & $(0,\alpha^{-\beta},\beta)$\tabularnewline

@@

Inverse chi-squared distribution & $(0,\infty)$ & $\frac{2^{-k/2}}{\Gamma(k/2)}x^{-k/2-1}e^{-1/(2x)}$ & $(-\frac{k}{2}-1,\frac{1}{2},-1)$\tabularnewline

@@

Inverse gamma distribution & $(0,\infty)$ & $\frac{\beta^{\alpha}}{\Gamma(\alpha)}x^{-\alpha-1}e^{-\beta/x}$ & $(-\alpha-1,\beta,-1)$\tabularnewline

@@

>>>>

\documentclass[thesis.tex]{subfiles}
% \documentclass{article}
% \csname remappendix\endcsname

@@

% \author{%
%   David S.~Hippocampus\thanks{Use footnote for providing further information
%     about author (webpage, alternative address)---\emph{not} for acknowledging
%     funding agencies.} \\
%   Department of Computer Science\\
%   Cranberry-Lemon University\\
%   Pittsburgh, PA 15213 \\
%   \texttt{hippo@cs.cranberry-lemon.edu} \\
%   % examples of more authors
%   % \And
%   % Coauthor \\
%   % Affiliation \\
%   % Address \\
%   % \texttt{email} \\
%   % \AND
%   % Coauthor \\
%   % Affiliation \\
%   % Address \\
%   % \texttt{email} \\
%   % \And
%   % Coauthor \\
%   % Affiliation \\
%   % Address \\
%   % \texttt{email} \\
%   % \And
%   % Coauthor \\
%   % Affiliation \\
%   % Address \\
%   % \texttt{email} \\
% }

% \sisetup{detect-all = true}

@@

\begin{table}[t]
    \centering
    \bgroup
{\small
\begin{tabular}{|M{0.15\linewidth} | p{0.75\linewidth}|}
\hline
\centering \multirow{2}{*}{\textbf{Ordering}} & 
$\max\{(\nu_1,\sigma_1,\rho_1),(\nu_2,\sigma_2,\rho_2)\}$ \\ & $\qquad \equiv \begin{cases}
(\nu_1,\sigma_1,\rho_1) & \text{ if } \limsup_{x\to\infty} \frac{x^\nu_1 e^{-\sigma_1 x^{\rho_1}}}{x^\nu_2 e^{-\sigma_2 x^{\rho_2}}} < +\infty \\
(\nu_2,\sigma_2,\rho_2) & \text{ otherwise.}
\end{cases}$ \\ \hline
\centering \multirow{2}{*}{\textbf{Addition}} &  $(\nu_{1},\sigma_{1},\rho_{1})\oplus(\nu_{2},\sigma_{2},\rho_{2})$ \\
 & $\equiv \begin{cases}
\max\{(\nu_{1},\sigma_{1},\rho_{1}),(\nu_{2},\sigma_{2},\rho_{2})\} & \text{ if }\rho_{1}\neq\rho_{2}\text{ or }\rho_{1},\rho_{2}<1\\
\left(\nu_{1}+\nu_{2}+1,\min\{\sigma_{1},\sigma_{2}\},1\right) & \text{ if }\rho_{1}=\rho_{2}=1\\
(\nu_{1}+\nu_{2}+\frac{2-\rho}{2},(\sigma_{1}^{-\frac{1}{\rho-1}}+\sigma_{2}^{-\frac{1}{\rho-1}})^{1-\rho},\rho) & \text{ if }\rho=\rho_{1}=\rho_{2}>1.
\end{cases}$ 
\\ \hline
\centering \textbf{Powers} & $(\nu,\sigma,\rho)^\beta \equiv (\frac{\nu+1}{\beta} - 1,\sigma,\frac{\rho}{\beta})$ for $\beta > 0$ \\\hline
\centering \textbf{Reciprocal*} & $(\nu,\sigma,\rho)^{-1} \equiv \begin{cases} (-\nu-2,\sigma,-\rho) & \text{ if } (\nu + 1)/\rho > 0 \text{ and } \rho \neq 0 \\
\mathcal{R}_2 & \text{ otherwise}
\end{cases}$ \\ \hline
\centering \textbf{Scalar \linebreak Multiplication} & $c(\nu,\sigma,\rho) \equiv (\nu,\sigma/|c|^\rho,\rho)$\\\hline
\multirow{3}{*}{\centering \textbf{Multiplication}} & $(\nu_{1},\sigma_{1},\rho_{1})\otimes(\nu_{2},\sigma_{2},\rho_{2})$ \\
&$\qquad \equiv\begin{cases}
\left(\frac{1}{\mu}\left(\frac{\nu_{1}}{|\rho_{1}|}+\frac{\nu_{2}}{|\rho_{2}|}+\frac{1}{2}\right),\sigma,-\frac{1}{\mu}\right) & \text{ if }\rho_{1},\rho_{2}<0\\
\left(\frac{1}{\mu}\left(\frac{\nu_{1}}{\rho_{1}}+\frac{\nu_{2}}{\rho_{2}}-\frac{1}{2}\right),\sigma,\frac{1}{\mu}\right) & \text{ if }\rho_{1},\rho_{2}>0\\
\mathcal{R}_{|\nu_1|} & \mbox{ if }\rho_{1}\leq0,\rho_{2}>0 \\
\mathcal{R}_{\min\{|\nu_1|,|\nu_2|\}} & \mbox{ if }\rho_{1}=0,\rho_{2}=0
\end{cases}$ \\
&where $\mu=\frac{1}{|\rho_{1}|}+\frac{1}{|\rho_{2}|}=\frac{|\rho_{1}|+|\rho_{2}|}{|\rho_{1}\rho_{2}|}$, $\sigma=\mu(\sigma_{1}|\rho_{1}|)^{\frac{1}{\mu|\rho_{1}|}}(\sigma_{2}|\rho_{2}|)^{\frac{1}{\mu|\rho_{2}|}}$.\\ \hline
\centering \textbf{Product of Densities} & $(\nu_1,\sigma_1,\rho_1)\&(\nu_2,\sigma_2,\rho_2) 
\equiv \begin{cases}
(\nu_{1}+\nu_{2},\sigma_{1},\rho_{1}) & \text{ if }\rho_{1}<\rho_{2}\\
(\nu_{1}+\nu_{2},\sigma_{1}+\sigma_{2},\rho) & \text{ if }\rho=\rho_{1}=\rho_{2}\\
(\nu_{1}+\nu_{2},\sigma_{2},\rho_{2}) & \text{ otherwise.}
\end{cases}$ \\ \hline
\centering \textbf{Functions ($L$-Lipschitz)} & 
$f(X_1,\dots,X_n) \equiv L \max\{X_1,\dots,X_n\}$ \\ \hline
    \end{tabular}
}
    \vspace{.25cm}
    \caption{\label{tab:gga_operations}Operations on random variables (e.g.$X_1 + X_2$) are 
    viewed as actions on density functions
    (e.g. convolution $(\nu_1, \sigma_1, \rho_1)\oplus (\nu_2, \sigma_2, \rho_2)$) and the tail parameters of the result are analyzed and reported.}
\egroup
\end{table}

@@

\textbf{Variational Inference} The optimization objective is the ELBO
\[
    E_q \log \frac{p(X)}{q(X)} \approx \frac{1}{N} \sum_{i=1}^N \log \frac{p(x_i)}{q(x_i)},\quad x_i \sim q
\]
Here, the density $p$ must also be evaluated so for simplicity
experiments in \cref{tab:vi} use closed-form marginalized densities for targets.
The overall trends also show that GGA yields consistent improvements as measured by both ELBO and importance-weighted
estimates of marginal likelihood and that the difference was greater when the tails of $p(z)$ were heavier.
The $\hat{k}$ diagnostics \citep{yao2018yes} corroborate our findings that variational inference succeeds
($\hat{k} < -1.2$) when a GGA with appropriately matched tails is used and fails ($\hat{k} > 1$) when
Gaussian tails are erroneously imposed.

@@

\begin{figure}[h]
	\centering
	\includegraphics[width=0.4\textwidth]{figures/LogNormal_Bulk.png}
	\includegraphics[width=0.4\textwidth]{figures/LogNormal_Tail.png}
	\caption{Estimation of the log-normal density by tail algebra applied to $V_k$.}
	\label{fig:lognormal}
\end{figure}

@@

Consider the toy example of a Normal target. This case is trivial for
Gaussian based methods and is oftentimes the initialization.
This lack of approximation gap in ADVI is seen in \Cref{fig:normal},
where we also see that GGA achieves similar approximation quality.
This is unsurprising as the GGA approximation in \Cref{tab:dist_list}
is also a Normal distribution.

\begin{figure}
	\centering
	\includegraphics[width=1.0\textwidth]{figures/de_normal_sum.png}
	\includegraphics[width=.4\textwidth]{figures/vi_normal_sum.png}
	\caption{Density estimation and VI against a known normal target}
	\label{fig:normal}
\end{figure}

@@

Now let $X_{ij} \sim N(0,1)$ and consider $tr X^\top X$.
Such quantities arise in the analysis of random projections.
It is important here to recognize that the power operation $X \mapsto X^2$ is not equivalent to the multiplication operation $X \mapsto X \otimes X$, as multiplication assumes independence.
\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{figures/jl-chi2.png}
	\caption{$5000$ samples of JL matrix trace (blue) vs GGA prediction (yellow)}
	\label{fig:jl_chi2}
\end{figure}

@@

{\small

@@

Gen. hyperbolic distribution & $(-\infty,\infty)$ & $\propto e^{\beta(x-\mu)}\frac{K_{\lambda-1/2}(\alpha\sqrt{\delta^{2}+(x-\mu)^{2}})}{(\delta^{2}+(x-\mu)^{2})^{1/4-\lambda/2}}$ & $(\lambda-1,\alpha-\beta,1)$\tabularnewline

@@

Gen. Normal distribution & $(-\infty,\infty)$ & $\frac{\beta}{2\alpha\Gamma(1/\beta)}\exp\left(-\left(\frac{|x-\mu|}{\alpha}\right)^{\beta}\right)$ & $(0,\alpha^{-\beta},\beta)$\tabularnewline

@@

Inv. chi-squared distribution & $(0,\infty)$ & $\frac{2^{-k/2}}{\Gamma(k/2)}x^{-k/2-1}e^{-1/(2x)}$ & $(-\frac{k}{2}-1,\frac{1}{2},-1)$\tabularnewline

@@

Inv. gamma distribution & $(0,\infty)$ & $\frac{\beta^{\alpha}}{\Gamma(\alpha)}x^{-\alpha-1}e^{-\beta/x}$ & $(-\alpha-1,\beta,-1)$\tabularnewline

@@

}
